Generating Empirically Optimized Composed Matrix Kernles from MATLAB Prototypes

### 2008-12-12 doc/papers/iccs09/paper.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@385 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{html}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{amsmath}

@@

% first the title is needed
\title{Generating Empirically Optimized Numerical Software from MATLAB Prototypes}

@@

        % the name(s) of the author(s) follow(s) next
        %
        % NB: Chinese authors should write their first names(s) in front of
        % their surnames. This ensures that the names appear correctly in
        % the running heads and the author index.
        %
\author{Elizabeth Jessup\inst{1} 
  \and Boyana Norris\inst{2}
  \and Albert Hartono\inst{3}
  \and Jeremy Siek\inst{1}}

@@

%\toctitle{Lecture Notes in Computer Science}
\tocauthor{Elizabeth Jessup, Boyana Norris, Albert Hartono, and Jeremy Siek}
\maketitle

\begin{abstract}
% between 75 and 150 words
The growing demand for higher levels of detail and accuracy in results means that the size and complexity of scientific computations is increasing at least as fast as the improvements in processor technology. Programming scientific applications is hard, and optimizing them for high performance is even harder.  The development of optimized codes requires extensive knowledge, not only of the costs of floating-point arithmetic but also of  memory access issues and compiler optimizations. Experiments show that the complexity of this hardware-software system means that performance is difficult to predict fully.  Therefore, computational scientists are often forced to choose between investing too much time in tuning code or  accepting performance that is significantly lower than the best achievable performance on a given architecture.
In this paper, we describe the first steps toward a fully automated system for the optimization of the matrix algebra kernels that are a foundational part of many scientific applications.  To generate highly optimized code from a high-level MATLAB prototype, we define a three-step approach.  To begin, we have developed a compiler that converts a MATLAB script into simple C code.   We then use the polyhedral optimization system PLuTo to optimize that code for coarse-grained parallelism and locality simultaneously. Finally, we annotate the resulting code with performance tuning directives and use the empirical performance tuning system Orio to generate many tuned versions of the same operation using different optimization techniques, such as loop unrolling and memory alignment. Orio performs an automated empirical search to select the best among the multiple optimized code variants. 
We discuss performance results showing that the code generated by using our system significantly outperforms not only the original simple C code but also code based on source BLAS, ATLAS-optimized BLAS, and Intel MKL routines.
\keywords{Code generation, empirical performance tuning}
\end{abstract}

@@

Historically, the research community has pursued two separate paths toward the goal of making software run at near-peak levels.  The first of these builds on research into compilers and their associated technologies.  The holy grail of compilation research is to take an arbitrary code as input and produce optimal code as output for a given language and hardware platform.  Despite the immense effort that has been poured into that approach, its success has been limited both by practical constraints (users will not tolerate compile times that extend into several days) and by the amount of detailed information the compiler can obtain about the software to be compiled and the hardware on which it is to execute~\cite{Ken99,Wol89,Wol96}.  The second path is to identify kernel routines that constitute the dominant costs of a wide variety of applications.  When such kernels can be identified and agreed upon by the members of the community, programmers with the required level of technical knowledge can concentrate on producing these optimized kernel libraries for architectures of interest. A prime example is the high-performance BLAS libraries produced by a combination of hardware vendors, independent software vendors, and researchers.  Developers who write their codes calling these routines can achieve high performance across all supported architectures. The efforts in high-performance library development aptly demonstrate, however, that the research into compiler technologies has not been fully successful.  If compilers could take the relatively simple code even for basic tasks such as a dense matrix-matrix product and produce optimal code, this second approach would be unnecessary.    Like compiler optimization, the kernel library-oriented approach has significant limitations.  Support for a variety of architectures is the most serious problem because the optimizations necessary to achieve near-peak performance are by nature not portable.   Lack of portability is particularly problematic because processor designs are constantly changing,  and constant updates are nearly impossible to maintain by hand.  It is a sad fact of the computing industry that by the time highly-optimized code is available for a given architecture, that architecture is generally well on its way to obsolescence.

We are working on a combination of the two approaches designed to overcome some of their shortcomings.   This paper describes our multipronged approach toward the development of software infrastructure for \emph{generating} automatically tuned libraries for matrix algebra computations.  We briefly overview relevant prior and current research efforts in Section~\ref{sec:background}. In Section~\ref{sec:approach} we describe our MATLAB to C compiler and the empirical performance tuning system Orio and its use in conjunction with the PLuTo tool suite to generate and empirically evaluate many tuned versions of the C code generated by the MATLAB compiler.  In Section~\ref{sec:experiments}, we provide performance results showing that the code generated by using our system significantly outperforms not only the original simple C code but also code based on source BLAS, ATLAS-optimized BLAS, and Intel MKL routines.
%Text in intro repeats abstract some, but that might not be a big deal for a conf paper.

@@

Existing optimizing MATLAB compilers, such as the MaJIC MATLAB compiler, include limited local optimizations for matrix expressions, but do not perform optimizations such as loop fusion across multiple operations as we do with the tools described in this paper. The telescoping languages project~\cite{telescopingurl,teleoverview,Ken99} uses techniques such as strength reduction, vectorization, and procedure specialization to optimize MATLAB scripts, but do not generate reusable optimized linear algebra routines as described in this paper.

@@

Specialized code generators circumvent the high costs of manual code generation . They include tools for basic dense linear algebra operations (ATLAS~\cite{WN147}, PhiPAC~\cite{bilmes97optimizing,phipacwww}), sparse linear algebra (OSKI~\cite{vuduc05}), FFTs~\cite{FFTW,Spiral}), stencil-based operations~\cite{kamil06}, and tensor computations (TCE~\cite{TCE}) among others.  It is often impossible to predict the performance of code precisely on modern computer architectures. Thus, many of these specialized code generators exploit search strategies to identify the best (or nearly best) code for a particular choice of problem parameters and machine. Such autotuning tools are not general but rather each is valuable in one specific problem domain. While not all applications can be written in terms of available libraries, the performance achieved with autotuning tools does convincingly demonstrate the advantage of source-to-source transformation over optimizing compilers.

The best performance today is obtained through approaches like that taken by PERI~\cite{PERI} which involves a labor intensive rewrite of portion of the code by a group of performance tools experts in collaboration with application scientists. Maintaining performance of the code as the application and hardware evolves means that the tuning process may need to be repeated. Our efforts aim to capture this performance tuning expertise in a way that can be integrated into the application. We will thus enable performance portability and ensure that the improvements to the application are long-lasting.

A number of source or binary transformation tools for performance-improving optimizations exist. LoopTool~\cite{LoopTool}, developed at Rice University, supports annotation-based loop fusion, unroll-and-jam, skewing and tiling.  A relatively new tool, POET~\cite{POET} also supports a number of loop transformations. POET offers a complex template-based syntax for defining transformations in a language-independent manner. Other recent research efforts whose goal, at least in part, is to enable optimizations of source code to be augmented with performance-related information include the X language~\cite{XLanguage} (a macro C-like language for annotating C code), the Broadway~\cite{broadway} compiler, and various meta-programming techniques~\cite{veldhuizen95,weise93,kiczales91,chiba95}. PLuTo~\cite{Pluto,uday08cc} is a source-to-source transformation tool for optimizing sequences of nested loops. PLuTo employs a polyhedral model of nested loops, where the dynamic instance (iteration) of each statement is viewed as an integer point in a well-defined space, called the statement's polyhedron. Combined with a characterization of data dependences, this representation allows the construction of mathematically correct complex loop transformations. The transformations target both improved cache locality and parallelism.

\section{Our Approach to Optimizing Composed BLAS Operations}
\label{sec:approach}

%\section{Optimizing Composed Operations}
%\label{sec:composed}

@@

The authors and other researchers have observed that it can be much more efficient to call a single specialized routine that performs multiple operations than to make successive calls to separate BLAS routines ~\cite{Ashby:uq,Blackford:2002vn,baker03blgmres,baker03lgmres,Dennis:2005tg,Howell:2008,gropp01,Vuduc:2003kl}.  Single routines that carry out more than one linear algebra operation are known as \emph{composed} BLAS. As an example, consider the pair of matrix-vector products $q  = Ap$, $s  = A^Tr$, where $A$ is a matrix and $p$, $q$, $r$, and $s$ are vectors, that comprise the computational
bottlenecks of the BiConjugate Gradient Method~\cite{Barrett:1994kx,Saad:2003fk}  and of the GEMVER kernel examined in Section~\ref{sec:matlab}.   These two operations can be implemented as a pair of calls to the BLAS routine {\tt GEMV} or they can be rewritten as a composed BLAS consisting of the doubly nested loop encompassing both matrix-vector multiplies. In the former case, the matrix $A$ is accessed twice, while in the latter it is accessed only once. Our preliminary results in prior research and in this paper indicate that this loop fusion leads to a routine that delivers significantly better performance than does a pair of calls to the best optimized BLAS {\tt GEMV} routines for large matrix orders.
Composed routines are the focus of the work presented here, but much of what we discuss generalizes to a much broader array of computations.

% While large speedups are probable, composition can  also force memory accesses to be consecutive, and it can

%lead to the need to bring more data into cache. These bad

% effects can be remedied by combining composition with other tuning techniques like cache blocking and software pipelining.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.7\textwidth]{figures/process.png}
\caption{Overview of the code generation and tuning process, optional steps are indicated with dashed arrows.}
\label{fig:process}
\end{figure}
To generate highly optimized code from a MATLAB prototype of the composed BLAS operation, we follow a three-step approach, illustrated in Figure~\ref{fig:process}.  To begin, we have developed a compiler that converts a MATLAB script into simple C code.
%~\cite{Siek}. 
After generating the C code from the high-level MATLAB prototype, we (optionally) use the source-to-source automatic parallelization tool PLuTo\cite{Pluto} %developed at Ohio State University
to optimize for coarse-grained parallelism and locality simultaneously. Using the results of the PLuTo analysis, we insert annotations into the C code, which are then processed by our extensible annotation system Orio to generate many tuned versions of the same operation using different optimization parameters. Orio then performs an empirical search to select the best among the multiple optimized code variants.

@@

\subsection{A MATLAB Compiler}
\label{sec:matlab}

In this section we describe the design and operation of the MATLAB to C compiler we have developed.    Figure~\ref{fig:compiler} gives an overview of the compilation process. The MATLAB kernel specification is parsed into a high-level intermediate representation in the form of a dataflow graph.  This dataflow graph is then iteratively processed until all of the implementation choices have been made.  The compilation process consists of three phases: analysis, lowering, and optimization, which are together iterated until all of the implementation decisions have been made.  The graph is then translated into C code.

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/compile.png}

\caption{Overview of the MATLAB compiler.}
\label{fig:compiler}
\end{figure}

\subsubsection{The Dataflow Graph}

An example dataflow graph for the GEMVER kernel, defined below, is given in figure~\ref{fig:gemver-dataflow}.
\begin{align*}
  A &\gets A + u_1 v_1^T + u_2 v_2^T \\[-0.5ex]
  x &\gets \beta A^T y + z \\[-0.5ex]
  w &\gets \alpha A x
\end{align*}
Each node represents a parameter of the kernel or an operation.  The arrows indicate the flow of data. At first the graph specifies what operations are to be performed but does not contain any implementation details. The symbol $\times$ in the depicted graph stands variously for outer product, matrix-vector multiplication, and scalar-vector multiplication, and it does not yet specify, for example, whether the outer products compute a row or a column at a time of the result matrix.

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/gemver-dataflow.png}

\caption{Dataflow graph for the GEMVER kernel.}
\label{fig:gemver-dataflow}
\end{figure}

\subsubsection{Analyze the Dataflow Graph}

@@

\subsection{Refine the Dataflow Graph}
\label{sec:refine}

@@

\subsubsection{Optimize the Dataflow Graph}

@@

\subsubsection{Translate the Dataflow Graph to C}

@@

%~\cite{Norris:2007}
Orio is an empirical tuning tool that takes annotated C code as input,
generates multiple transformed versions of the annotated code, and
empirically evaluates the performance of the generated codes,
ultimately choosing the best-performing version to use in production
runs.

@@

Finally, while Orio currently generates various loop transformations, it does not generate calls to existing library implementations. For this paper, we implemented the BLAS versions of all operations manually. Our past success with identifying Blue Gene/L intrinsics from expression graphs motivate further research on identifying BLAS and other established library operations from the expression graphs of the MATLAB code.

%\singlespacing
%% or stuff it in a footnote
%\section*{Acknowledgments}
%This work was supported by the Mathematical, Information, and
%Computational Sciences Division subprogram of the Office of Advanced
%Scientific Computing Research, Office of Science, U.S. Dept. of Energy,
%under Contract DE-AC02-06CH11357. 
%TODO: need Colorado and Ohio acknowledgments

>>>>

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{html}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amsmath}

@@

% first the title is needed
\title{Generating Empirically Optimized Composed Matrix Kernles from MATLAB Prototypes}

@@

\author{Elizabeth Jessup\inst{1} 
  \and Boyana Norris\inst{2}
  \and Albert Hartono\inst{3}
  \and Jeremy Siek\inst{1}}

@@

\tocauthor{Elizabeth Jessup, Boyana Norris, Albert Hartono, and Jeremy Siek}
\maketitle

\input{abstract}

@@

Historically, the research community has pursued two separate paths toward
the goal of making software run at near-peak levels.  The first path builds
on research into compilers and their associated technologies.  One of the
main goals of compilation research is to take an arbitrary code as input and
produce optimal code as output for a given language and hardware platform.
The success of this approach has been limited by a number of factors: (i)
optimal mappings between the computation graph and the hardware are expensive
(often NP-complete) to compute; (ii) potentially useful information that
could aid optimization cannot be represented in general-purpose languages
such as C and Fortran; and (iii) user control of compiler optimizations is
limited and varies from compiler to compiler, and (iv) apart from differences
in execution time, it is difficult for users to evaluate the effectiveness of
different compiler optimizations.

The second path to performance optimization is to identify kernel
routines that dominate the execution time of a wide variety of
applications.  
%When such kernels can be identified and agreed upon by the
%members of the community, programmers with the required level of technical
%knowledge can concentrate on producing these optimized kernel libraries for
%architectures of interest. 
An example is the high-performance BLAS libraries produced by a combination
of hardware vendors, independent software vendors, and researchers.
Developers who write their codes calling these routines can achieve high
performance across all supported architectures. The efforts in
high-performance library development aptly demonstrate, however, that the
research into compiler technologies has not been fully successful.  If
compilers could take the relatively simple code such as a dense matrix-matrix
product and produce optimal code, this second approach would be unnecessary.
Like compiler optimization, the kernel library-oriented approach has
significant limitations.  Support for a variety of architectures is the most
serious problem because the optimizations necessary to achieve near-peak
performance are by nature not portable, requiring manual update or
reimplementation for each new architecture.

This paper describes a combination of the two approaches designed to overcome
some of their shortcomings. We take a multipronged approach
toward the development of software infrastructure for \emph{generating}
automatically tuned libraries for matrix algebra computations.  We briefly
overview relevant prior and current research efforts in
Section~\ref{sec:background}. In Section~\ref{sec:approach} we describe our
MATLAB~\cite{matlab:webpage} to C compiler and the empirical performance tuning system Orio and its
use in conjunction with the PLuTo tool suite to generate and empirically
evaluate many tuned versions of the C code generated by the MATLAB compiler.
In Section~\ref{sec:experiments}, we provide performance results showing that
the code generated by using our system significantly outperforms not only the
original simple C code but also code based on source BLAS, ATLAS-optimized
BLAS, and Intel MKL routines.
%Text in intro repeats abstract some, but that might not be a big deal for a conf paper.

@@

Existing optimizing MATLAB compilers, such as the MaJIC MATLAB compiler~\cite{MaJIC},
include limited local optimizations for matrix expressions, but do not
perform optimizations such as loop fusion across multiple operations as we do
with the tools described in this paper. The telescoping languages
project~\cite{telescopingurl,teleoverview,Ken99} uses techniques such as
strength reduction, vectorization, and procedure specialization to optimize
MATLAB scripts, but do not generate reusable optimized linear algebra
routines as described in this paper.

@@

Specialized code generators circumvent the high costs of manual code
generation . They include tools for basic dense linear algebra operations
(ATLAS~\cite{WN147}, PhiPAC~\cite{bilmes97optimizing,phipacwww}), sparse
linear algebra (OSKI~\cite{vuduc05}), FFTs~\cite{FFTW,Spiral}), stencil-based
operations~\cite{kamil06}, and tensor computations (TCE~\cite{TCE}) among
others.  While these libraries target a specific kernel, our approach aims 
at enabling the definition of arbitrary kernels involving dense matrix linear algebra.
%It is often impossible to predict the performance of code precisely
%on modern computer architectures. Thus, many of these specialized code
%generators exploit search strategies to identify the best (or nearly best)
%code for a particular choice of problem parameters and machine. Such
%autotuning tools are not general but rather each is valuable in one specific
%problem domain. 
%While not all applications can be written in terms of available libraries,
%the performance achieved with autotuning tools does convincingly demonstrate
%the benefits of automated source code transformations.

%The best performance today is obtained through approaches like that taken by
%PERI~\cite{PERI} which involves a labor intensive rewrite of portion of the
%code by a group of performance tools experts in collaboration with
%application scientists. Maintaining performance of the code as the
%application and hardware evolves means that the tuning process may need to be
%repeated. Our efforts aim to capture this performance tuning expertise in a
%way that can be integrated into the application. We will thus enable
%performance portability and ensure that the improvements to the application
%are long-lasting.

A number of source or binary transformation tools for general
performance-improving optimizations exist. LoopTool~\cite{LoopTool},
developed at Rice University, supports annotation-based loop fusion,
unroll-and-jam, skewing and tiling.  A relatively new tool, POET~\cite{POET}
also supports a number of loop transformations. POET offers a complex
template-based syntax for defining transformations in a language-independent
manner. Other recent research efforts whose goal, at least in part, is to
enable optimizations of source code to be augmented with performance-related
information include the X language~\cite{XLanguage} (a macro C-like language
for annotating C code), the Broadway~\cite{broadway} compiler, and various
meta-programming
techniques~\cite{veldhuizen95,weise93,kiczales91,chiba95}. PLuTo~\cite{Pluto,uday08cc}
is a source-to-source transformation tool for optimizing sequences of nested
loops. PLuTo employs a polyhedral model of nested loops, where the dynamic
instance (iteration) of each statement is viewed as an integer point in a
well-defined space, called the statement's polyhedron. Combined with a
characterization of data dependences, this representation allows the
construction of mathematically correct complex loop transformations. The
transformations target both improved cache locality and parallelism.

\section{Optimizing Composed BLAS Operations}
\label{sec:approach}

@@

The authors and other researchers have observed that it can be much more
efficient to call a single specialized routine that performs multiple
operations than to make successive calls to separate BLAS routines (e.g., see
~\cite{Blackford:2002vn,baker03lgmres,,gropp01,Vuduc:2003kl}.
%~\cite{Ashby:uq,Blackford:2002vn,baker03blgmres,baker03lgmres,Dennis:2005tg,Howell:2008,gropp01,Vuduc:2003kl}.
Single routines that carry out more than one linear algebra operation are
known as \emph{composed} BLAS. As an example, consider the pair of
matrix-vector products $q = Ap$, $s = A^Tr$, where $A$ is a matrix and $p$,
$q$, $r$, and $s$ are vectors, that comprise the computational bottlenecks of
the BiConjugate Gradient Method~\cite{Barrett:1994kx,Saad:2003fk} and of the
GEMVER kernel examined in Section~\ref{sec:matlab}.  These two operations can
be implemented as a pair of calls to the BLAS routine {\tt GEMV} or they can
be rewritten as a composed BLAS consisting of the doubly nested loop
encompassing both matrix-vector products. In the former case, the matrix
$A$ is accessed twice, while in the latter it is accessed only once. Our
preliminary results in prior research and in this paper indicate that 
loop fusion leads to a routine that delivers significantly better performance
than does a pair of calls to the best optimized BLAS {\tt GEMV} routines for
large matrix orders.  Composed routines are the focus of the work presented
here, but much of what we discuss generalizes to a much broader array of
computations.

To generate highly optimized code from a MATLAB prototype of the composed BLAS operation, we follow a three-step approach, illustrated in Figure~\ref{fig:process}.  To begin,
%
\begin{wrapfigure}{r}{3in}
\vspace{-.2in}
\centering
\includegraphics[width=3in]{figures/process.png}
\caption{Overview of the code generation and tuning process, optional steps are indicated with dashed arrows.}
\label{fig:process}
\vspace{-.1in}
\end{wrapfigure}
%
%~\cite{Siek}.  
we have developed a compiler that converts a MATLAB script into simple C code.
After generating the C code from the high-level MATLAB prototype, we (optionally) use the source-to-source automatic parallelization tool PLuTo\cite{Pluto} %developed at Ohio State University
to optimize for coarse-grained parallelism and locality simultaneously. Using the results of the PLuTo analysis, we insert annotations into the C code, which are then processed by our extensible annotation system Orio to generate many tuned versions of the same operation using different optimization parameters. Orio then performs an empirical search to select the best among the multiple optimized code variants.

@@

\vspace{-.2in}
\subsection{A MATLAB Compiler}
\label{sec:matlab}
%\input{matlab}
In this section we describe the design and operation of the MATLAB to C compiler we have developed~\cite{Siek}.    Figure~\ref{fig:compiler} gives an overview of the compilation process. 
%
\begin{wrapfigure}{r}{2in}
\vspace{-.2in}
\centering
\includegraphics[width=2in]{figures/compile.png}
\caption{The MATLAB compiler.}
\label{fig:compiler}
\vspace{-.3in}
\end{wrapfigure}
The MATLAB kernel specification is parsed into a high-level intermediate representation in the form of a dataflow graph.  This dataflow graph is then iteratively processed until all of the implementation choices have been made.  The compilation process consists of three phases: analysis, lowering, and optimization, which are together iterated until all of the implementation decisions have been made.  The graph is then translated into C code.

\subsubsection{The Dataflow Graph.}

An example dataflow graph for the GEMVER kernel, defined below, is given in figure~\ref{fig:gemver-dataflow}.
\begin{figure}[ht]
\begin{minipage}{.3\textwidth}
\begin{align*}
  A &\gets A + u_1 v_1^T + u_2 v_2^T \\[-0.5ex]
  x &\gets \beta A^T y + z \\[-0.5ex]
  w &\gets \alpha A x
\end{align*} 
\end{minipage}
\begin{minipage}{.6\textwidth}
\includegraphics[width=3in]{figures/gemver-dataflow.png}
\end{minipage}
\caption{Dataflow graph (right) for the GEMVER kernel (left).}
\label{fig:gemver-dataflow}
\end{figure}

Each node represents a parameter of the kernel or an operation.  The arrows 
indicate the flow of data. At first the graph specifies what operations are to be performed but does not contain any implementation details. The symbol $\times$ in the depicted graph stands variously for outer product, matrix-vector multiplication, and scalar-vector multiplication, and it does not yet specify, for example, whether the outer products compute a row or a column at a time of the result matrix.

\subsubsection{Analyze the Dataflow Graph.}

@@

\subsubsection{Refine the Dataflow Graph.}
\label{sec:refine}

@@

\subsubsection{Optimize the Dataflow Graph.}

@@

\subsubsection{Translate the Dataflow Graph to C.}

@@

%~\cite{Norris:2007}
Orio\cite{Norris:2007,ipdps09} is an empirical tuning tool that takes annotated C code as input,
generates multiple transformed versions of the annotated code, and
empirically evaluates the performance of the generated codes,
ultimately choosing the best-performing version to use in production
runs.

@@

Finally, while Orio currently generates various loop transformations, it does not generate calls to existing library implementations. For this paper, we implemented the BLAS versions of all operations manually. Our past success with identifying Blue Gene/L intrinsics from expression graphs motivate further research on identifying BLAS and other established library operations from the expression graphs of the MATLAB code.
\cite{Siek}
%\singlespacing
%% or stuff it in a footnote
%\section*{Acknowledgments}
%This work was supported by the Mathematical, Information, and
%Computational Sciences Division subprogram of the Office of Advanced
%Scientific Computing Research, Office of Science, U.S. Dept. of Energy,
%under Contract DE-AC02-06CH11357. 
%TODO: need Colorado and Ohio acknowledgments
